Aldebaran~b's temperate past uncovered in planet search data

### 2018-02-05 paper/aldebaran.tex
celerite to appendix

Fortunately, the damped and driven harmonic oscillator GP can be written as the
solution of a class of stochastic ordinary differential equations, the
Continuous Auto-Regressive Moving Average (CARMA) models, whose likelihoods can
be evaluated in linear time. For problems such as this, we therefore have access
to these powerful computational tools for inferring their power spectral
densities.  Our treatment of CARMA models is described more fully in Appendix
\ref{carmatheory}.   \citet{Kelly2014} also give more details on CARMA
processes in an astronomical context, and show how a state-space model of the
process $y(t)$ can be tracked through a time-series of uncertain observations
$y_k = y\left( t_k \right) + \epsilon_k$, with Gaussian noise $\epsilon_k$ of
zero mean and known (heteroskedastic) variance, using a Kalman filter to produce
a likelihood function $p\left( y_k \mid \sigma, r_i, b_j \right)$ depending on
the amplitude of the stochastic forcing, $\sigma$, the (possibly complex)
eigenfrequencies of the ODE, $r_i$, and parameters describing the power spectrum
of the stochastic forcing, $b_j$ in $\mathcal{O}\left( n \right)$ computational
time.  Additionally, \citet{Foreman-Mackey2017} show how a dimensional expansion
can be used to reduce the standard GP covariance matrix for this
same process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
observations at times $t_k$, to a banded form that can then be
Cholesky-decomposed in the standard GP likelihood function in
$\mathcal{O}\left( n \right)$ time.

@@

>>>>

Fortunately, the damped and driven harmonic oscillator GP can be written as the
solution of a class of stochastic ordinary differential equations, the
Continuous Auto-Regressive Moving Average (CARMA) models, whose likelihoods can
be evaluated in linear time. For problems such as this, we therefore have access
to these powerful computational tools for inferring their power spectral
densities.  Our treatment of CARMA models is described more fully in Appendix
\ref{carmatheory}.   \citet{Kelly2014} also give more details on CARMA
processes in an astronomical context, and show how a state-space model of the
process $y(t)$ can be tracked through a time-series of uncertain observations
$y_k = y\left( t_k \right) + \epsilon_k$, with Gaussian noise $\epsilon_k$ of
zero mean and known (heteroskedastic) variance, using a Kalman filter to produce
a likelihood function $p\left( y_k \mid \sigma, r_i, b_j \right)$ depending on
the amplitude of the stochastic forcing, $\sigma$, the (possibly complex)
eigenfrequencies of the ODE, $r_i$, and parameters describing the power spectrum
of the stochastic forcing, $b_j$ in $\mathcal{O}\left( n \right)$ computational
time.

@@

The Kalman filter implementation of the CARMA likelihood is not unique. 
\citet{Foreman-Mackey2017} show how a dimensional expansion
can be used to reduce the standard GP covariance matrix for this
same process,  with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
observations at times $t_k$, to a banded form that can then be
Cholesky-decomposed in the standard GP likelihood function in
$\mathcal{O}\left( n \right)$ time. These `\emph{celerite}' models are formally equivalent to the Kalman CARMA model and differ only in their computational implementation.
